---
title: "Effect of Neonic Pesticides to Honey Production"
output: html_document
---

## 1. INTRODUCTION

  Neonicotinoids or neonics are a class of insecticides used in modern crop protection that specifically targets the nicotinic acetylcholine receptors of insects (Jeschke and Nauen, 2008). Neonics have been in the market since the 90s and remain to be the most widely-used pesticide in the USA and China (Muth and Leonard, 2019). While the selectivity of the product aims to improve crop health without causing adverse effects to non-target organisms such as humans, some inevitable consequences arise nevertheless.

  Bees are especially important in ecology for pollination and floral diversity. The decline in bee population has been correlated with the decline in pollinator-reliant out-crossing plant species (Biesmeijer et al., 2006). This decline in bee population has been implicated in the use of neonics, although reports show that acute exposure to neonics only affects foraging performance and not the survival or mortality of bees (Muth and Leonard, 2019).
  
  In this report, we intend to investigate whether the use of neonics indeed affect the total production of honey in four regions of the US (midwest, west, northeast, and south) using the dataset "Honeybees and Neonic Pesticides" from the US Department of Agriculture (USDA) and the US Geological Survery (USGS). Total production, the product of average yield per colony and the number of colonies in the region, is used as an indirect measurement of bee population in the assumption that higher total honey production corresponds to higher bee population.
  
  At the end of this data analysis, we aim to analyze the relationship between the total production of honey with use of Neonic pesticides in four USA regions; compare the effect of different types of neonics, specifically clothianidin, imidacloprid, thiamethoxam, acetamiprid, and thaicloprid, on total honey production; and identify the trend, if any exists, in total honey production from 1991 to 2017.


## 2. DATA

  The dataset was accessed online from URL: https://www.kaggle.com/kevinzmith/honey-with-neonic-pesticide, entitled "Honeybees and Neonic Pesticides". Variables involving honey production are from USDA, namely, number of honey-producing colonies, honey yield per colony, total production, stocks, price per lb, and production value. 
  
  On the other hand, data concerning use of neonic pesticides from all over US are from USGS; the variables include total and individual amounts of each neonic type (clothianidin, imidacloprid, thiamethoxam, acetamiprid, thaicloprid). Neonic data is collected by country, state, region, and year. Kevin Zmith classified the state into four regions: South, West, Midwest, and Northeast using US Census.

```{r}
#LIBRARIES USED
library(tidyverse)
library(dplyr)
library(ggplot2)
library(ggpubr)
library(Rmisc)
library(rstatix)
```
  
```{r}
#HONEY AND NEONIC PESTICIDES DATA
honey <- read.csv("./Honey_v03.csv")
head(honey)
glimpse(honey)
```


## 3. DATA ANALYSIS

  In the following data analysis, changes in total production are evaluated by region. This analysis focuses on factors that affect honey production as an indication of the bee population. Interactions between regions are beyond the limits of this study, since we treated regions independently. Lastly, since the USDA and USGS data are obtained separately, some missing values limited the analysis to visual graphs instead of statistical tests (see 3.3). Furthermore, due to the diversity of factors not taken into account and included in the dataset, further investigation must be performed to understand substantially how the use of neonics contribute to the decline in the bee population.

# 3.1. Total Neonic Production

  We first evaluated the effect of neonic pesticides on total honey production. The total honey production takes into account the number of colonies and the honey yield per colony (lb). Here, we analyzed the correlation of total honey production as a function of total neonic pesticides (kg) per region.

```{r}
byproduction <- honey %>%
  group_by(Region) 

Figure1 <- ggplot(data = byproduction, mapping = aes(x = log(nAllNeonic), 
  y = log(totalprod), group = Region)) +
  facet_grid(.~ Region) + geom_point() + 
  geom_smooth(method = "lm", se = FALSE) +
  labs (x = "log of Total Neonic (kg)", 
        y = "log of Total Production (lb)", 
        title = " Bee Honey Production under Neonic Pesticide by Region")

Figure1
```

  The relationship of the total production and total neonic is positive for all regions except in the midwest (Figure1). To identify the strength and direction of the relationship between the two variables in different regions, we used the Spearman correlation test. Indeed, out of the four regions, the correlation of total honey production and neonic pesticides in the midwest is weak, negative, and not statistically significant (p > 0.05, rho= -0.103). On the other hand, total honey production and total neonic in the south, west, and northeast regions show weak, positive, and statistically significant correlations (Table1). While the linear relationships may be weak, there is statistical evidence to prove that total production of honey is positively correlated with total neonic in all regions except the midwest. Why this relationship exists cannot be explained by data present in the dataset. Other factors such as geographic attributes including population and local consumption of neonics should be taken into account in future studies.


```{r}
#SPEARMAN CORRELATION BY REGION
neonicstatmw <- honey %>%
  filter(Region == "Midwest")
cortest.mw <- cor.test(neonicstatmw$totalprod, neonicstatmw$nAllNeonic, method = "spearman", exact = FALSE)

neonicstatwest <- honey %>%
  filter(Region == "West")
cortest.w <- cor.test(neonicstatwest$totalprod, neonicstatwest$nAllNeonic, method = "spearman", exact = FALSE)

neonicstatne <- honey %>%
  filter(Region == "Northeast")
cortest.ne <- cor.test(neonicstatne$totalprod, neonicstatne$nAllNeonic, method = "spearman", exact = FALSE)

neonicstatsouth <- honey %>%
  filter(Region == "South")
cortest.s <- cor.test(neonicstatsouth$totalprod, neonicstatsouth$nAllNeonic, method = "spearman", exact = FALSE)

#FRAMING THE STATISTICAL RESULTS
Region <- c("Midwest", "West", "Northeast", "South")
p.value <- c(cortest.mw$p.value, cortest.w$p.value, 
             cortest.mw$p.value, cortest.s$p.value)
rho <- c(-0.1027632, 0.3549792, 0.3510381, 0.4034019)
significance <- c("ns", "*", "*", "*")

allneonic <- cbind(Region, p.value, rho, significance)
Table1 <- data.frame(allneonic); Table1
```

# 3.2 Type of Neonic

  According to the USGS data, there are fives types of neonics used in the US States, namely, clothianidin, imidacloprid, thiamethoxam, acetamiprid, thaicloprid. The relationships of total honey production and the amount of neonic by type are shown below (Figure2).

```{r}
byneonic <- select(honey, totalprod, 12:16, Region)
byneonic <- na.omit(byneonic)
byneonicA <- gather(byneonic, "Type", "Amount", 2:6)
byneonicB <- byneonicA %>%
  mutate(logtotalprod = log(totalprod), logamount = log(Amount))


Figure2 <- ggplot(byneonicB, aes(x = logamount, y = logtotalprod, color = Type)) + 
  geom_point() + facet_grid(.~ Region) + 
   labs(title = "Effects of Different Neonic Types to Honey Production by Region", 
        x = "log of Amount of Neonic (kg)", 
        y = "log of Total Honey Production (lb)") +
   scale_fill_discrete(name = "Type", labels = c("Acetamiprid", "Clothianidin", 
                                                 "Imidacloprid", "Thiacloprid", 
                                                 "Thiamethoxam"))  +
  geom_smooth(method = lm, se = FALSE); Figure2
```

  All neonic types show a weak correlation with total honey production. Of all, only thiamethoxam and clothianidin show significant correlation across all regions, although some are positive while others are negative. We cannot be certain of the possible reasons for these variations, but we propose two hypotheses for further testing. 
  The first hypothesis involves testing for the lethality of all neonics. Neonics were significantly found to have dose-dependent detrimental effects on foraging performance of bees (Muth and Leonard, 2019). Data on mortality rate due to acute exposure of bees to neonics should be investigated.
  Second, we can test for the total contribution of each neonic type to the total neonic recorded in 3.1. The significance levels may have been affected by the non-uniformity of the data; some neonic types were present at high levels, while some are not. Furthermore, the USDS dataset have several missing data, for which our analysis cannot fully compensate.

```{r}
#FILTER DATASET BY REGION
midwest <- byneonic %>%
  filter(Region == "Midwest")
northeast <- byneonic %>%
  filter(Region == "Northeast")
south <- byneonic %>%
  filter(Region == "South")
west <- byneonic %>%
  filter(Region == "West")

<<<<<<< HEAD
#SPEARMAN CORRELATION OF NEONIC TYPE AND TOTALPROD BY REGION
#Midwest
acemid.cor <- cor.test(midwest$totalprod, midwest$Acetamiprid, 
                       method = "spearman", exact = FALSE)
clomid.cor <- cor.test(midwest$totalprod, midwest$Clothianidin, 
                       method = "spearman", exact = FALSE)
midmid.cor <- cor.test(midwest$totalprod, midwest$Midacloprid, 
                       method = "spearman", exact = FALSE)
thiamethid.cor <- cor.test(midwest$totalprod, midwest$Thiamethoxam, 
                           method = "spearman", exact = FALSE)
thiaclomid.cor <- cor.test(midwest$totalprod, midwest$Thiacloprid, 
                           method = "spearman", exact = FALSE)

#Northeast
acene.cor <- cor.test(northeast$totalprod, northeast$Acetamiprid, 
                      method = "spearman", exact = FALSE)
clone.cor <- cor.test(northeast$totalprod, northeast$Clothianidin, 
                      method = "spearman", exact = FALSE)
midne.cor <- cor.test(northeast$totalprod, northeast$Midacloprid, 
                      method = "spearman", exact = FALSE)
thiamethne.cor <- cor.test(northeast$totalprod, northeast$Thiamethoxam, 
                           method = "spearman", exact = FALSE)
thiaclone.cor <- cor.test(northeast$totalprod, northeast$Thiacloprid, 
                          method = "spearman", exact = FALSE)

#South
acesouth.cor <- cor.test(south$totalprod, south$Acetamiprid, 
                         method = "spearman", exact = FALSE)
closouth.cor <- cor.test(south$totalprod, south$Clothianidin, 
                         method = "spearman", exact = FALSE)
midsouth.cor <- cor.test(south$totalprod, south$Midacloprid, 
                         method = "spearman", exact = FALSE)
thiamethsouth.cor <- cor.test(south$totalprod, south$Thiamethoxam, 
                              method = "spearman", exact = FALSE)
thiaclosouth.cor <- cor.test(south$totalprod, south$Thiacloprid,
                             method = "spearman", exact = FALSE)

#West
acewest.cor <- cor.test(west$totalprod, west$Acetamiprid, 
                        method = "spearman", exact = FALSE)
clowest.cor <- cor.test(west$totalprod, west$Clothianidin, 
                        method = "spearman", exact = FALSE)
midwest.cor <- cor.test(west$totalprod, west$Midacloprid, 
                        method = "spearman", exact = FALSE)
thiamethwest.cor <- cor.test(west$totalprod, west$Thiamethoxam, 
                             method = "spearman", exact = FALSE)
thiaclowest.cor <- cor.test(west$totalprod, west$Thiacloprid, 
                            method = "spearman", exact = FALSE)

#FRAMING THE STATISTICAL RESULTS
Region. <- c("Midwest", "Midwest", "Midwest", "Midwest", "Midwest", 
             "Northeast", "Northeast", "Northeast", "Northeast","Northeast",
             "South", "South", "South", "South", "South", "West", "West", 
             "West", "West","West")
NeonicType <- c("Acetamiprid", "Clothianidin", "Midacloprid", "Thiamethoxam", 
                "Thiacloprid", "Acetamiprid", "Clothianidin", "Midacloprid", 
                "Thiamethoxam", "Thiacloprid", "Acetamiprid", "Clothianidin", 
                "Midacloprid", "Thiamethoxam", "Thiacloprid", "Acetamiprid", 
                "Clothianidin", "Midacloprid", "Thiamethoxam", "Thiacloprid")
p.value. <- c(acemid.cor$p.value, clomid.cor$p.value, midmid.cor$p.value, 
              thiamethid.cor$p.value, thiaclomid.cor$p.value, acene.cor$p.value, 
              clone.cor$p.value, midne.cor$p.value, thiamethne.cor$p.value, 
              thiaclone.cor$p.value, acesouth.cor$p.value, closouth.cor$p.value, 
              midsouth.cor$p.value, thiamethsouth.cor$p.value, 
              thiaclosouth.cor$p.value, acewest.cor$p.value, 
              clowest.cor$p.value, midwest.cor$p.value, 
              thiamethsouth.cor$p.value, thiaclowest.cor$p.value)
rho. <- c(-0.3406599, -0.178460 ,-0.04370683 ,-0.1223294 ,-0.2207321  ,0.178998
=======
#Correlation test per neonic type per region

acemid.cor <- cor.test(midwest$totalprod, midwest$Acetamiprid, method = "spearman", exact = FALSE)
acemid.cor

clomid.cor <- cor.test(midwest$totalprod, midwest$Clothianidin, method = "spearman", exact = FALSE)
clomid.cor


midmid.cor <- cor.test(midwest$totalprod, midwest$Imidacloprid, method = "spearman", exact = FALSE)
midmid.cor


thiamethid.cor <- cor.test(midwest$totalprod, midwest$Thiamethoxam, method = "spearman", exact = FALSE)
thiamethid.cor

thiaclomid.cor <- cor.test(midwest$totalprod, midwest$Thiacloprid, method = "spearman", exact = FALSE)
thiaclomid.cor


#northeast

acene.cor <- cor.test(northeast$totalprod, northeast$Acetamiprid, method = "spearman", exact = FALSE)
acene.cor

clone.cor <- cor.test(northeast$totalprod, northeast$Clothianidin, method = "spearman", exact = FALSE)
clone.cor


midne.cor <- cor.test(northeast$totalprod, northeast$Imidacloprid, method = "spearman", exact = FALSE)
midne.cor


thiamethne.cor <- cor.test(northeast$totalprod, northeast$Thiamethoxam, method = "spearman", exact = FALSE)
thiamethne.cor


thiaclone.cor <- cor.test(northeast$totalprod, northeast$Thiacloprid, method = "spearman", exact = FALSE)
thiaclone.cor

#south

acesouth.cor <- cor.test(south$totalprod, south$Acetamiprid, method = "spearman", exact = FALSE)
acesouth.cor

closouth.cor <- cor.test(south$totalprod, south$Clothianidin, method = "spearman", exact = FALSE)
closouth.cor


midsouth.cor <- cor.test(south$totalprod, south$Imidacloprid, method = "spearman", exact = FALSE)
midsouth.cor


thiamethsouth.cor <- cor.test(south$totalprod, south$Thiamethoxam, method = "spearman", exact = FALSE)
thiamethsouth.cor


thiaclosouth.cor <- cor.test(south$totalprod, south$Thiacloprid, method = "spearman", exact = FALSE)
thiaclosouth.cor

#west

acewest.cor <- cor.test(west$totalprod, west$Acetamiprid, method = "spearman", exact = FALSE)
acewest.cor

clowest.cor <- cor.test(west$totalprod, west$Clothianidin, method = "spearman", exact = FALSE)
clowest.cor


midwest.cor <- cor.test(west$totalprod, west$Imidacloprid, method = "spearman", exact = FALSE)
midwest.cor


thiamethwest.cor <- cor.test(west$totalprod, west$Thiamethoxam, method = "spearman", exact = FALSE)
thiamethwest.cor


thiaclowest.cor <- cor.test(west$totalprod, west$Thiacloprid, method = "spearman", exact = FALSE)
thiaclowest.cor

#making dataframe of the significance of statistical results
Region <- c("Midwest", "Midwest", "Midwest", "Midwest", "Midwest", "Northeast", "Northeast", "Northeast", "Northeast","Northeast","South", "South", "South", "South", "South", "West", "West", "West", "West","West")
NeonicType <- c("Acetamiprid", "Clothianidin", "Imidacloprid", "Thiamethoxam", "Thiacloprid", "Acetamiprid", "Clothianidin", "Imidacloprid", "Thiamethoxam", "Thiacloprid", "Acetamiprid", "Clothianidin", "Imidacloprid", "Thiamethoxam", "Thiacloprid", "Acetamiprid", "Clothianidin", "Imidacloprid", "Thiamethoxam", "Thiacloprid")
p.value <- c(acemid.cor$p.value, clomid.cor$p.value, midmid.cor$p.value, thiamethid.cor$p.value, thiaclomid.cor$p.value, acene.cor$p.value, clone.cor$p.value, midne.cor$p.value, thiamethne.cor$p.value, thiaclone.cor$p.value, acesouth.cor$p.value, closouth.cor$p.value, midsouth.cor$p.value, thiamethsouth.cor$p.value, thiaclosouth.cor$p.value, acewest.cor$p.value, clowest.cor$p.value, midwest.cor$p.value, thiamethsouth.cor$p.value, thiaclowest.cor$p.value)

rho <- c(-0.3406599, -0.178460 ,-0.04370683 ,-0.1223294 ,-0.2207321  ,0.178998
>>>>>>> 2a3fad0b5cf978cde0008ab9fe462b6844638c1a
         ,0.1504854 ,0.2958545 ,0.178656 ,0.2972762, 0.06931884, -0.03242255,
         0.5089368, 0.1946103, -0.2852726, 0.06372726, -0.07198862 , 0.3833301,
         0.1946103, 0.1439576)
significance. <- c("*", "*", "ns", "*", "*", "ns", "*", "ns", "*", "ns","ns", 
                   "*", "*", "*", "ns", "ns", "*", "*", "*")
  
neonic.type.df <- cbind(Region., NeonicType, p.value., rho., significance.)
Table2 <- as.data.frame(neonic.type.df); Table2
```


# 3.3. Year

  Total production of honey throughout 1991 to 2017 is visualized for four regions (Figure 3). The midwest shows the highest production throughout the years, while Northeast shows the lowest. Statistical analysis for any significant difference between years was impossible due to the absence of data from select states in each region which prevents pairwise comparison of repeated measures. However, despite the growing industry of neonic production starting in 2003, total production of honey from bee colonies have been fluctuating in the midwest and the west even before. Furthermore, northeastern and southern total production have been relatively steady for years. 
  This could suggest that neonic production does not necessarily induce an effect on total production of honey throughout the years. Other factors must be considered why production in the northeast and the south are maintained, while there is unsteady decline in the midwest and the west.
  Interestingly, the midwest remains the highest producer of neonics followed by the south. Why the midwest still remains the highest producer of honey despite the high production of neonics could perhaps be explained by biological and geographical factors which are beyond the scope of the study.
  Without the use of statistical tests to point out the significance of differences in total production of honey throughout the years, we cannot conclude definitely whether the use of neonics has a relationship with total production from 1991 to 2017, although graphic analysis proves it unlikely.

```{r}
byyear <- select(honey, year, Region, totalprod)
byyear$totalprod <- log(byyear$totalprod)
byyearSE <- summarySE(byyear, measurevar = "totalprod", groupvars = c("Region", "year"))
pd <- position_dodge(0.1)

yearplot <- ggplot(byyearSE, aes(x = year, y = totalprod, color = "red", group = 1)) + 
    geom_errorbar(aes(ymin = byyearSE$totalprod - ci, ymax = byyearSE$totalprod + ci), 
                  width=.1, position = pd, alpha = 0.4) +
    geom_line(position = pd) +
    geom_point(position = pd, size = 2, shape = 21, fill = "white") +
  facet_wrap(.~Region) + 
  labs(title = "A. Total Production", 
       subtitle = "1991 - 2017", 
       y = "log(Total Honey Production)",
       x = "Year") +
  theme(legend.position = "none")

neonicyear <- ggplot(honey, aes(x = year, y = nAllNeonic)) + geom_col() + 
  facet_wrap(.~ Region) + labs(title = "B. Neonic Production",
                               subtitle = "1991 - 2017",
                               y = "Total Neonic Production",
                               x = "Year")

Figure3 <- ggarrange(yearplot, neonicyear, ncol = 2); Figure3
```

## 4. SUMMARY

The relationship between the use of neonic pesticides in US agricultural lands
and honey production of bees varies in different regions. Particularly, the
correlation of the two variables are significant for all regions, except for
the Midwest (p > 0.05, rho= -0.103). The total honey production were also
analyzed in terms of each of the five neonic types. We found the presence
of two neonics, thiamethoxam and clothianidin, to be statistically significant
in all regions. The trend of honey production and use of neonics from 1991
to 2017 reveal that neonic production do not necessarily affect the decline or
increase of honey production.

It should be noted that the dataset is a combination of separate datasets, one
from USDS and the other from USDA. We are not certain if they involved evaluating
the same set of bee colonies. Hence, we recommend the future researchers to 
assess the use of neonic pesticides and honey production in the same location
to ensure higher chances of interaction between the variables. We also suggest
taking note of other biological and geological events that can possibly affect
the decline in bee honey production.

## REFERENCES

Muth, F. and Leonard, A. S. (2019) ‘A neonicotinoid pesticide impairs foraging, but not learning, in free-flying bumblebees’, Scientific Reports, 9(1), p. 4764. doi: 10.1038/s41598-019-39701-5.

Jeschke, P. and Nauen, R. (2008) ‘Neonicotinoids—from zero to hero in insecticide chemistry’, Pest Management Science, 64(11), pp. 1084–1098. doi: 10.1002/ps.1631.

Biesmeijer, J. C. et al. (2006) ‘Parallel Declines in Pollinators and Insect-Pollinated Plants in Britain and the Netherlands’, Science, 313(5785), pp. 351–354. doi: 10.1126/science.1127863.