---
title: "Capstone Markdown"
output: html_document
---
<<<<<<< HEAD

```{r}
library(dplyr)
library(tidyverse)
library(ggplot2)
```

PLOT 1. neonic vs. total.prod (color = region)
```{r}
honey.csv <- read.csv("Honey_v03.csv")

neonicvtotalprod <- honey.csv %>%
  group_by(Region) 

ggplot(data = neonicvtotalprod, mapping = aes(x = nAllNeonic, y = totalprod, col = Region)) +
  geom_point() +
  labs (x = "Total Neonic (kg)", y = "Total Production (lb)", title = " Bee Honey Production under Neonic Pesticide by Region")
```


PLOT 2. neonic type vs. total prod (color = region)



```{r}
neonictype <- honey.csv %>%
  gather(key = c(nCLOTHIANIDIN, nIMIDACLOPRID, nTHIAMETHOXAM, nACETAMIPRID, nTHIACLOPRID), value = value) %>%
  print()

neonictype <- honey.csv %>%
  pivot_longer(cols = c(12,13,14,15,16), names_to = "neonictype", values_to = "x")


a <- ggplot(data = neonicvtotalprod, mapping = aes(x = nCLOTHIANIDIN, y = totalprod, col = Region)) + geom_point() + 
  labs(x = "Clothianidin (kg)", y = "Total Production (lb)")

b <- ggplot(data = neonicvtotalprod, mapping = aes(x = nIMIDACLOPRID, y = totalprod, col = Region)) + geom_point () +
  labs(x = "Imidacloprid (kg)", y = "Total Production (lb)")

c <-  ggplot(data = neonicvtotalprod, mapping = aes(x = nTHIAMETHOXAM, y = totalprod, col = Region)) + geom_point() +
  labs(x = "Thiamethoxam (kg)", y = "Total Production (lb)")

d <- ggplot(data = neonicvtotalprod, mapping = aes(x = nACETAMIPRID, y = totalprod, col = Region)) + geom_point() +
  labs(x = "Acetamiprid (kg)", y = "Total Production (lb)")

e <- ggplot(data = neonicvtotalprod, mapping = aes(x = nTHIACLOPRID, y = totalprod, col = Region)) + geom_point() +
  labs(x = "Thiacloprid (kg)", y = "Total Production (lb)")

library(cowplot)
library(ggpubr)
library(grid)

ggarrange(a, b, c, d, e + rremove("x.text"),
ncol = 2, nrow = 3) #+ grid_arrange() +
  #labs(x = "Neocin Type") #not a facetgrid pero similar na pero looking for ways na ma han ay sha like
# one label for y-axis
# only one area for the legend
# put general labels for title, x, and y - axis
```


=======
LIBRARIES
```{r setup, include=FALSE}
library(tidyverse)
install.packages("ggubr")
library(ggpubr)
install.packages("ggubr")
library(rstatix)
```

1. TEST FOR NORMALITY

shapiro.test(data$column)

if: p > 0.05, data is normal
    p < 0.05, data is non normal
    
2. IF TEST IS NORMAL, TEST FOR HOMOGENEITY

Perform this test only if the parametric test requires homogenous data

var.test(y variable ~ x variable, data = data) ---> requires data transformation

3. PERFORM STATISTICAL TEST

Bitaw bata, search lang unsaon ang data sa kada stat imo gamiton. Naay uban required nga
itransform ang data. If naa kay questions example unsa nga stat gamiton if categorical vs continuous
or whatever pwede ra ka mochat or visit ni:

https://www.scribbr.com/statistics/statistical-tests/

After nimo read ani pwede ra nimo ni iremove
>>>>>>> 019134c01a9ed8f9b77baa1887aa40cf5035e6c1


