---
title: "Capstone Markdown"
output: html_document
---

```{r}
library(dplyr)
library(tidyverse)
library(ggplot2)
library(stats)
```

PLOT 1. neonic vs. total.prod (color = region)
```{r}
honey.csv <- read.csv("Honey_v03.csv")


neonicvtotalprod <- honey.csv %>%
  group_by(Region) 


ggplot(data = neonicvtotalprod, mapping = aes(x = nAllNeonic, y = log(totalprod), group = Region)) +
  facet_grid(.~ Region) +
  geom_point() +
  geom_smooth(method = "lm", se = FALSE) +
  labs (x = "Total Neonic (kg)", y = "Total Production (lb)", title = " Bee Honey Production under Neonic Pesticide by Region")



#plot1.mod <- lm(data = honey.csv, formula = log(totalprod) ~ nAllNeonic)
#summary(plot1.mod)

#Since there is a significant relationship in total production between
#the four regions (insert p-value), we perform post-hoc analysis

# library(multcomp)


#post_hoc_honey <- glht()


# STATISTICAL TEST, CORRELATION PER REGION
neonicstatmw <- honey %>%
  filter(Region == "Midwest")
cormw <- cor.test(neonicstatmw$totalprod, neonicstatmw$nAllNeonic, method = "spearman", exact = FALSE)
cormw

neonicstatwest <- honey %>%
  filter(Region == "West")
corw <- cor.test(neonicstatwest$totalprod, neonicstatwest$nAllNeonic, method = "spearman", exact = FALSE)
corw

neonicstatne <- honey %>%
  filter(Region == "Northeast")
corne <- cor.test(neonicstatne$totalprod, neonicstatne$nAllNeonic, method = "spearman", exact = FALSE)
corne

neonicstatsouth <- honey %>%
  filter(Region == "South")
cors <- cor.test(neonicstatsouth$totalprod, neonicstatsouth$nAllNeonic, method = "spearman", exact = FALSE)
cors

# MAKING A DATAFRAME OF STATISTICAL RESULTS
Region <- c("Midwest", "West", "Northeast", "South")
p.value <- c(cormw$p.value, corw$p.value, corne$p.value, cors$p.value)
rho <- c(-0.1027632, 0.3549792, 0.3510381, 0.4034019)
significance <- c("ns", "*", "*", "*")

allneonic <- cbind(Region, p.value, rho, significance)
allneonic <- data.frame(allneonic); allneonic
```

HELLO BATA, CHECK AKONG STAT GIBUHAT PARA DI NA KA MAGLISOD PERFORM SA IMOHA

PLOT 2. neonic type vs. total prod (color = region)


```{r}
byneonic <- select(honey.csv, totalprod, 12:16, Region); byneonic
byneonic <- na.omit(byneonic)
byneonicA <- gather(byneonic, "Type", "Amount", 2:6)
byneonicB <- byneonicA %>%
  mutate(logtotalprod = log(totalprod), logamount = log(Amount))
byneonicB

ggplot(byneonicB, aes(x = logamount, y = logtotalprod, color = Type)) + 
  geom_point() + facet_grid(.~ Region) + 
   labs(Title = "Effects of different Neonic types to honey production by Region", x = "Amount of Neonic (kg)", y = "Total Honey Production (lb)") +
   scale_fill_discrete(name = "Type", labels = c("Acetamiprid", "Clothianidin", "Imidacloprid", "Thiacloprid", "Thiamethoxam"))  +
  geom_smooth(method = lm, se = FALSE)

#filter dataset by region
midwest <- byneonic %>%
  filter(Region == "Midwest")

northeast <- byneonic %>%
  filter(Region == "Northeast")

south <- byneonic %>%
  filter(Region == "South")

west <- byneonic %>%
  filter(Region == "West")

#Correlation test per neocin type per region

acemid.cor <- cor.test(midwest$totalprod, midwest$Acetamiprid, method = "spearman", exact = FALSE)
acemid.cor

clomid.cor <- cor.test(midwest$totalprod, midwest$Clothianidin, method = "spearman", exact = FALSE)
clomid.cor


midmid.cor <- cor.test(midwest$totalprod, midwest$Midacloprid, method = "spearman", exact = FALSE)
midmid.cor


thiamethid.cor <- cor.test(midwest$totalprod, midwest$Thiamethoxam, method = "spearman", exact = FALSE)
thiamethid.cor

thiaclomid.cor <- cor.test(midwest$totalprod, midwest$Thiacloprid, method = "spearman", exact = FALSE)
thiaclomid.cor


#northeast

acene.cor <- cor.test(northeast$totalprod, northeast$Acetamiprid, method = "spearman", exact = FALSE)
acene.cor

clone.cor <- cor.test(northeast$totalprod, northeast$Clothianidin, method = "spearman", exact = FALSE)
clone.cor


midne.cor <- cor.test(northeast$totalprod, northeast$Midacloprid, method = "spearman", exact = FALSE)
midne.cor


thiamethne.cor <- cor.test(northeast$totalprod, northeast$Thiamethoxam, method = "spearman", exact = FALSE)
thiamethne.cor


thiaclone.cor <- cor.test(northeast$totalprod, northeast$Thiacloprid, method = "spearman", exact = FALSE)
thiaclone.cor

#south

acesouth.cor <- cor.test(south$totalprod, south$Acetamiprid, method = "spearman", exact = FALSE)
acesouth.cor

closouth.cor <- cor.test(south$totalprod, south$Clothianidin, method = "spearman", exact = FALSE)
closouth.cor


midsouth.cor <- cor.test(south$totalprod, south$Midacloprid, method = "spearman", exact = FALSE)
midsouth.cor


thiamethsouth.cor <- cor.test(south$totalprod, south$Thiamethoxam, method = "spearman", exact = FALSE)
thiamethsouth.cor


thiaclosouth.cor <- cor.test(south$totalprod, south$Thiacloprid, method = "spearman", exact = FALSE)
thiaclosouth.cor

#west

acewest.cor <- cor.test(west$totalprod, west$Acetamiprid, method = "spearman", exact = FALSE)
acewest.cor

clowest.cor <- cor.test(west$totalprod, west$Clothianidin, method = "spearman", exact = FALSE)
clowest.cor


midwest.cor <- cor.test(west$totalprod, west$Midacloprid, method = "spearman", exact = FALSE)
midwest.cor


thiamethwest.cor <- cor.test(west$totalprod, west$Thiamethoxam, method = "spearman", exact = FALSE)
thiamethwest.cor


thiaclowest.cor <- cor.test(west$totalprod, west$Thiacloprid, method = "spearman", exact = FALSE)
thiaclowest.cor

#making dataframe of the significance of statistical results
Region <- c("Midwest", "Midwest", "Midwest", "Midwest", "Midwest", "Northeast", "Northeast", "Northeast", "Northeast","Northeast","South", "South", "South", "South", "South", "West", "West", "West", "West","West")
NeonicType <- c("Acetamiprid", "Clothianidin", "Midacloprid", "Thiamethoxam", "Thiacloprid", "Acetamiprid", "Clothianidin", "Midacloprid", "Thiamethoxam", "Thiacloprid", "Acetamiprid", "Clothianidin", "Midacloprid", "Thiamethoxam", "Thiacloprid", "Acetamiprid", "Clothianidin", "Midacloprid", "Thiamethoxam", "Thiacloprid")
p.value <- c(acemid.cor$p.value, clomid.cor$p.value, midmid.cor$p.value, thiamethid.cor$p.value, thiaclomid.cor$p.value, acene.cor$p.value, clone.cor$p.value, midne.cor$p.value, thiamethne.cor$p.value, thiaclone.cor$p.value, acesouth.cor$p.value, closouth.cor$p.value, midsouth.cor$p.value, thiamethsouth.cor$p.value, thiaclosouth.cor$p.value, acewest.cor$p.value, clowest.cor$p.value, midwest.cor$p.value, thiamethsouth.cor$p.value, thiaclowest.cor$p.value)

rho <- c(-0.3406599, -0.178460 ,-0.04370683 ,-0.1223294 ,-0.2207321  ,0.178998
         ,0.1504854 ,0.2958545 ,0.178656 ,0.2972762, 0.06931884, -0.03242255,
         0.5089368, 0.1946103, -0.2852726, 0.06372726, -0.07198862 , 0.3833301,
         0.1946103, 0.1439576)

1.533013e-08 < 0.05

significance <- c("s", "s", "ns", "s", "s", "ns", "s", "ns", "s", "ns","ns", "s", "s", "s", "ns", "ns", "s", "s", "s")
  
neonic.type.df <- cbind(Region, NeonicType, p.value, rho, significance)
neonic.type.df <- as.data.frame(neonic.typee)
View((neonic.type.df))

```{r}
#stat:


```



LIBRARIES
```{r setup, include=FALSE}
library(tidyverse)
install.packages("ggubr")
library(ggpubr)
install.packages("ggubr")
library(rstatix)
```

1. TEST FOR NORMALITY

shapiro.test(data$column)

if: p > 0.05, data is normal
    p < 0.05, data is non normal
    
2. IF TEST IS NORMAL, TEST FOR HOMOGENEITY

Perform this test only if the parametric test requires homogenous data

var.test(y variable ~ x variable, data = data) ---> requires data transformation

3. PERFORM STATISTICAL TEST

Bitaw bata, search lang unsaon ang data sa kada stat imo gamiton. Naay uban required nga
itransform ang data. If naa kay questions example unsa nga stat gamiton if categorical vs continuous
or whatever pwede ra ka mochat or visit ni:

https://www.scribbr.com/statistics/statistical-tests/

After nimo read ani pwede ra nimo ni iremove


HAHAHA magic

IMPORT DATA
```{r}
honey <- read.csv("./Honey_v03.csv")
honey
```

YEAR VS. PROD
```{r}
year_plot <- ggplot(honey, aes(year, totalprod)) + geom_point()
year_plot
```