---
title: "Capstone Markdown"
output: html_document
---

```{r}
library(dplyr)
library(tidyverse)
library(ggplot2)
```

PLOT 1. neonic vs. total.prod (color = region)
```{r}
honey.csv <- read.csv("Honey_v03.csv")


neonicvtotalprod <- honey.csv %>%
  group_by(Region) 


ggplot(data = neonicvtotalprod, mapping = aes(x = nAllNeonic, y = log(totalprod), group = Region)) +
  facet_grid(.~ Region) +
  geom_point() +
  geom_smooth(method = "lm", se = FALSE) +
  labs (x = "Total Neonic (kg)", y = "Total Production (lb)", title = " Bee Honey Production under Neonic Pesticide by Region")



plot1.mod <- lm(data = honey.csv, formula = log(totalprod) ~ nAllNeonic)
summary(plot1.mod)

#Since there is a significant relationship in total production between
#the four regions (insert p-value), we perform post-hoc analysis

# library(multcomp)


#post_hoc_honey <- glht()


# nonparametric counterpart of anova for independent samples
#nonparametric post hoc test for unpaired or independent (usually mann whitney)
```


PLOT 2. neonic type vs. total prod (color = region)


```{r}
byneonic <- select(honey.csv, totalprod, 12:16, Region); byneonic
byneonic <- na.omit(byneonic)
byneonicA <- gather(byneonic, "Type", "Amount", 2:6)
byneonicB <- byneonicA %>%
  mutate(logtotalprod = log(totalprod), logamount = log(Amount))
byneonicB

ggplot(byneonicB, aes(x = logamount, y = logtotalprod, color = Type)) + 
  geom_point() + facet_grid(.~ Region) + 
   labs(Title = "Effects of different Neonic types to honey production by Region", x = "Amount of Neonic (kg)", y = "Total Honey Production (lb)") +
   scale_fill_discrete(name = "Type", labels = c("Acetamiprid", "Clothianidin", "Imidacloprid", "Thiacloprid", "Thiamethoxam"))  +
  geom_smooth(method = lm, se = FALSE)

#tagsatagsa ug lm sa mga lines (5 x 4 regions) --> to prove statistical significance 
 #use method = spearman
```

```{r}
#stat:


```



LIBRARIES
```{r setup, include=FALSE}
library(tidyverse)
install.packages("ggubr")
library(ggpubr)
install.packages("ggubr")
library(rstatix)
```

1. TEST FOR NORMALITY

shapiro.test(data$column)

if: p > 0.05, data is normal
    p < 0.05, data is non normal
    
2. IF TEST IS NORMAL, TEST FOR HOMOGENEITY

Perform this test only if the parametric test requires homogenous data

var.test(y variable ~ x variable, data = data) ---> requires data transformation

3. PERFORM STATISTICAL TEST

Bitaw bata, search lang unsaon ang data sa kada stat imo gamiton. Naay uban required nga
itransform ang data. If naa kay questions example unsa nga stat gamiton if categorical vs continuous
or whatever pwede ra ka mochat or visit ni:

https://www.scribbr.com/statistics/statistical-tests/

After nimo read ani pwede ra nimo ni iremove


HAHAHA magic

IMPORT DATA
```{r}
honey <- read.csv("./Honey_v03.csv")
honey
```

YEAR VS. PROD
```{r}
year_plot <- ggplot(honey, aes(year, totalprod)) + geom_point()
year_plot
```